---
title: "170A_Project"
author: "Shuyi Chen"
date: "2025-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/tc/Downloads/CHOC_Infusion')
```

```{r}
library(tidybayes)
library(lubridate)
library(tidyverse)
```


```{r}
abx_ord <- read.csv('uci_mds_ip_abx_ords.csv')
patients <- read.csv('uci_mds_ip_enc.csv')
pressor <- read.csv('uci_mds_ip_pressors_admins.csv')
vitals <- read_csv('uci_mds_ip_vitals.csv')
labs <- read_csv('uci_mds_ip_lab_results.csv')
```

```{r}
#View(abx_ord)
#View(patient)
summary(abx_ord)
```

```{r}
summary(patients)
```

```{r}
patients <- patients %>%
  mutate(REG_DT = as.POSIXct(REG_DT, format = "%m/%d/%y %H:%M"),
         DISCH_DT = as.POSIXct(DISCH_DT, format = "%m/%d/%y %H:%M"))

abx_ord <- abx_ord %>%
  mutate(ORD_DT = as.POSIXct(ORD_DT, format = "%m/%d/%y %H:%M"))


combined_data <- inner_join(abx_ord, patients, by = c("MRN_DI", "FIN_DI"))

```


```{r}
##Calculate Delays
combined_data <- combined_data %>%
  mutate(delay_minutes = as.numeric(difftime(ORD_DT, REG_DT, units = "mins")),
         delay_hours = delay_minutes / 60)

summary(combined_data$delay_hours)
```
```{r}
negative_delays_count <- combined_data %>%
  filter(delay_minutes < 0) %>%
  nrow()
negative_delays_count
```


```{r}
negative_delays <- combined_data %>% 
  filter(delay_minutes < 0)

earliest_visits <- patients %>%
  group_by(MRN_DI) %>%
  summarize(earliest_REG_DT = min(REG_DT, na.rm = TRUE))

negative_delays_prior <- negative_delays %>%
  left_join(earliest_visits, by = "MRN_DI") %>%
  filter(earliest_REG_DT < REG_DT)

total_unique_patients <- negative_delays %>% distinct(MRN_DI) %>% nrow()
unique_patients_with_prior <- negative_delays_prior %>% distinct(MRN_DI) %>% nrow()
proportion_prior <- unique_patients_with_prior / total_unique_patients

#proportion of patients from negative delays have prior visits
print(proportion_prior)
```

```{r}
unique_patients_overall <- combined_data %>% 
  distinct(MRN_DI) %>% 
  nrow()
print(unique_patients_overall)
```
```{r}
positive_delays <- combined_data %>% 
  filter(delay_minutes > 0)

# Plot the distribution of positive delay hours
ggplot(positive_delays, aes(x = delay_hours)) +
  geom_histogram(bins = 100, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Positive Delay Hours",
       x = "Delay Hours",
       y = "Number of Patients") +
  theme_minimal()


negative_delays <- combined_data %>% 
  filter(delay_minutes < 0)

# Plot the distribution of positive delay hours
ggplot(negative_delays, aes(x = delay_hours)) +
  geom_histogram(bins = 100, fill = "orange", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Positive Delay Hours",
       x = "Delay Hours",
       y = "Number of Patients") +
  theme_minimal()

```


```{r}
#observing only 0-48 hrs of delays
positive_delays_48 <- combined_data %>%
  filter(delay_minutes > 0, delay_hours <= 48)

nrow(positive_delays_48)
```


```{r}
summary(pressor)
```

```{r}
#unique patients in 0-48hrs delay group
 positive_delays_48 %>% 
  distinct(MRN_DI) %>% 
  nrow()
```
```{r}
#patients in both delays and labs
positive_delays_48 %>%
  semi_join(labs, by = "MRN_DI") %>%  
  distinct(MRN_DI) %>%
  nrow()
#patients in both delays and vitals
positive_delays_48 %>%
  semi_join(vitals, by = "MRN_DI") %>%  
  distinct(MRN_DI) %>%
  nrow()
#patients in both delays and pressor
positive_delays_48 %>%
  semi_join(pressor, by = "MRN_DI") %>%  
  distinct(MRN_DI) %>%
  nrow()
```

Since most patients appear to have lab results, and lab results contains multiple continuous variables that we don't have enough prior patient to compute, we will be dropping patients with no lab results. Also, because there are very few patients in pressor data, we are converting pressor data to binary variable. 

```{r}
#filter patients with no lab results
positive_delays_48_cleaned <- positive_delays_48 %>%
  semi_join(labs, by = c("MRN_DI", "FIN_DI"))

nrow(positive_delays_48_cleaned)
nrow(positive_delays_48)
```

We dropped about 6000 observations when filtering out patients with no lab results. 

```{r}
#encounters that have pressor data
pressor_encounters <- pressor %>%
  distinct(MRN_DI, FIN_DI) %>%
  mutate(pressor_use = 1)
nrow(pressor_encounters)
```
```{r}
#create binary variable indicating whether the patient encounter is on pressor or not
positive_delays_48_cleaned <- positive_delays_48_cleaned %>%
  left_join(pressor_encounters, by = c("MRN_DI", "FIN_DI")) %>%
  mutate(pressor_use = if_else(is.na(pressor_use), 0, 1))
```


```{r}
summary(positive_delays_48)
```

```{r}

```

